### 2021-02-01 [2020-08]_ACMTOMS_(MQSI)/paper/4-analysis.tex
Initial response to TOMS editor.

\heading{4. COMPLEXITY AND SENSITIVITY}
Algorithms 1, 2, and 4 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
the time, still ${\cal O}(n)$, is spent solving the banded linear
system of equations for the B-spline coefficients. Thus for $n$ data
points, the overall execution time is ${\cal O}(n)$. The quadratic
facet model produces a unique sensitivity to input perturbation, as
small changes in input may cause different quadratic facets to be
associated with a breakpoint, and thus different initial derivative
estimates can be produced. This phenomenon is depicted in Figure
1. Despite this sensitivity, the quadratic facet model is still
preferred because it generally provides aesthetically pleasing (low
{\it wiggle}) initial estimates of the first and second derivatives at
the breakpoints while perfectly capturing local low-order phenomena
(e.g., a run of points on a straight line) in the data. The binary
search for a point near the monotone boundary in $(\tau_1$, $\alpha$,
$\beta$, $\gamma)$ space is preferred because it results in monotone
quintic spline interpolants that are nearer to initial estimates than
a policy that strictly shrinks derivative values.

>>>>

\heading{4. COMPLEXITY AND SENSITIVITY}
Algorithms 1, 2, and 4 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
the time, still ${\cal O}(n)$, is spent solving the banded linear
system of equations for the B-spline coefficients. Thus for $n$ data
points, the overall execution time is ${\cal O}(n)$. The quadratic
facet model produces a unique sensitivity to input perturbation, as
small changes in input may cause different quadratic facets to be
associated with a breakpoint, and thus different initial derivative
estimates can be produced. This phenomenon is depicted in Figure
1. Despite this sensitivity, the quadratic facet model is still
preferred because it generally provides aesthetically pleasing (low
{\it wiggle}) initial estimates of the first and second derivatives at
the breakpoints while perfectly capturing local low-order phenomena
(e.g., a run of points on a straight line) in the data. The binary
search for a point on the monotone boundary in $(\tau_1$, $\alpha$,
$\beta$, $\gamma)$ space is preferred because it results in monotone
quintic spline interpolants that are nearer to initial estimates than
a policy that strictly shrinks derivative values.
